### 2012-11-13 doc/progress_plan/report.tex
Added section on programming language selection

Despite lacking behind the imagination of authors and script writers, over the last 60 years there
has been a huge amount of progress in AI techniques. The phrase `Artificial
Intelligence' was coined in 1956\cite{russell2010artificial} and can be used as
an umbrella term, describing many subfields from ``learning and perception
to... diagnosing diseases\it{(Ibid)}."

@@

Liakata et al. (2012) describe a system for automatically processing and
categorising sentences in a research paper according to their respective
CoreSC element\cite{citeulike:10444769}. SAPIENTA
(\url({http://www.sapientaproject.com}) was trained using the ART corpus and
can achieve promising accuracies when categorising sentences \it{(Ibid)}. With
the authors' permissions, Partridge will make use of SAPIENTA when introducing
new papers into its repository. It is hoped that preclassifying papers will
make the development of Partridge more manageable within the timeframe
available and will also allow more effort to be put into other classification
problems rather than re-implementing a SAPIENTA clone. Any modifications that
have to be made to SAPIENTA to integrate it into Partridge will be submitted
back to the original authors to help them improve their works.

After pre-processing by SAPIENTA is complete, Partridge will need to do some
classification work of its own to determine the area of study (i.e. Physics,
Chemistry, Biology), type of paper (Literature review, Case Study, Experiment)
and the polarity of the result (positive, negative, inconclusive). These tasks
involve implementing existing AI techniques and applying them specifically to
the task of text classification. To avoid the time consuming task of re-implementing these
methods, it was decided that a library should be used instead.

\subsubsection{NLTK}

There are several existing libraries to facilitate Natural Language Processing.
Many are written for Java \cite{mallet2002}\cite{cunningham2011text} and are
very complex or not well documented. The Natural Language Toolkit (NLTK) is a
simple and intuitive library written for Python. Bird(2009) states that NLTK
was designed ``to provide an intuitive framework along with substantial
building blocks, giving users a practical knowledge of NLP". The project is
relatively mature in comparison to the above named Java libraries. There is
also a free book that accompanies the project available at
\url{http://nltk.org/book/} which provides a huge amount of information on how
to implement many popular NLP techniques using the library. For this reason
NLTK was chosen as the primary accompanying library for the project.

@@

\subsubsection{Social Citation and Recommendation Engines}
Social citation and recommendation engines also provide a partial solution to
the `information overload' problem.  Services like Goodreads
(\url{http://www.goodreads.com/}) and CiteUlike
(\url{http://www.citeulike.org/}) allow you to register your interest in
specific authors and subjects. This allows the sites to build up a profile of
the sorts of materials that you might be interested in and provide lists of
recommendations as in Figure \cite{social_indexes}.

@@

These systems have the ability to make recommendations to the user without
requiring specific keywords or search terms. They are able to do this by
learning the user's profile and taking into account the preferences of their
'friends' and their browsing history. However, the above-named systems do not
take into account the content of the paper or book. They only deal with
metadata as can be seen in Figure \ref{fig:social_searches}. This means that
important discriminatory information that could be contained within the actual
document content is overlooked completely. Partridge will analyse documents on
a sentence-by-sentence and possibly word-by-word basis, thereby taking account
of any embedded information that could be missed by these social metadata
systems.

@@

@@

>>>>

Despite lacking behind the imagination of authors and script writers, over the last 60 years there
has been a huge amount of progress in AI techniques. The phrase `Artificial
Intelligence' was coined in 1956\cite{russell2010artificial} and can be used as
an umbrella term, describing many subfields from ``learning and perception
to... diagnosing diseases\emph{(Ibid)}."

@@

Liakata et al. (2012) describe a system for automatically processing and
categorising sentences in a research paper according to their respective
CoreSC element\cite{citeulike:10444769}. SAPIENTA
(\url{http://www.sapientaproject.com}) was trained using the ART corpus and
can achieve promising accuracies when categorising sentences (\emph{Ibid}). With
the authors' permissions, Partridge will make use of SAPIENTA when introducing
new papers into its repository. It is hoped that preclassifying papers will
make the development of Partridge more manageable within the timeframe
available and will also allow more effort to be put into other classification
problems rather than re-implementing a SAPIENTA clone. Any modifications that
have to be made to SAPIENTA to integrate it into Partridge will be submitted
back to the original authors to help them improve their works.

After pre-processing by SAPIENTA is complete, Partridge will need to do some
classification work of its own to determine the area of study (i.e. Physics,
Chemistry, Biology), type of paper (Literature review, Case Study, Experiment)
and the polarity of the result (positive, negative, inconclusive). These tasks
involve implementing existing AI techniques and applying them specifically to
the task of text classification. To avoid the time consuming task of re-implementing these
methods, it was decided that a library should be used instead. Different NLP
libraries and their properies are discussed below in Section
\ref{sec:libschoice}.

@@

\subsubsection{Social Citation and Recommendation Engines}
Social citation and recommendation engines also provide a partial solution to
the `information overload' problem.  Services like Goodreads
(\url{http://www.goodreads.com/}) and CiteUlike
(\url{http://www.citeulike.org/}) allow you to register your interest in
specific authors and subjects. This allows the sites to build up a profile of
the sorts of materials that you might be interested in and provide lists of
recommendations as in Figure \ref{social_indexes}.

@@

These systems have the ability to make recommendations to the user without
requiring specific keywords or search terms. They do this by learning the
user's profile and taking into account the preferences of their 'friends' and
their browsing history. However, the above-named systems do not take into
account the content of the paper or book. They only deal with metadata as can
be seen in Figure \ref{fig:social_searches}. This means that important
discriminatory information that could be contained within the actual document
content is overlooked completely. Partridge will analyse documents on a
sentence-by-sentence and possibly word-by-word basis, thereby taking account of
any embedded information that could be missed by these social metadata systems.

@@

Partridge is based upon several fairly complex and new ideas in AI and NLP.
For the project to be a success it is imperative that the right language,
environment and toolkits are used to avoid unnecessary work on existing
technologies and allow priority to be given to the novel aspects of the
project.

The language that Partridge will be written in is one of the most fundamental
and important choices to be made. The language choice effects not only the end
performance of the application, but the speed at which the project is developed
and the portability of the end code \cite{britton2008}. To avoid a learning
overhead, it was decided that the programming language should be one that the
author is familiar with. This reduced the choice of language down to C, Java
and Python.

The C programming language was invented by Kernighan and Ritchie and
published in 1978\cite{ritchie1978c}. The language is small and
optimised\cite{prinz2002c} and therefore compiled C programs tend to run
extremely fast. Unfortunately, C is designed to be general and lack restriction
\cite{ricthie1978c} which is often an advantage for programmers who favour
optimised code over excessive error checking. However, this means that
debugging C programms can be quite long winded and challenging. Since the NLP
aspects of Partridge present an amibitious challenge in themselves, having to
debug applications without help from a managed programming environment is not
desirable.

The Java programming language on the other hand, does provide memory management
and error checking\cite{Coffey2008} at the cost of program performance. Java is
a pseudo-compiled language that is translated into an intermediate bytecode
which is then interpreted at runtime by an interpreter on the client computer
called a Virtuam Machine(VM). Gosling (2000) describes java as as a
``general-purpose...  programming language [that allows] developers to write a
program once and then be able to run it everywhere on the
internet\cite{gosling2000java}." Java provides a solid and stable programming
environment and stringent error checking, both of which would be advantageous
in the development of Partridge.  The language syntax is very verbose and a lot
of `boiler-plate' code must be written before a program can be executed.
Therefore Java is not suitable as a rapid prototyping language. Since
Partridge's development requires a steep learning curve for the author, a
language that allows quick prototyping with minimal programming.

Python is a programming language with a famously shallow learning curve,
readable syntax and dynamic inspection that facilitates rapid prototyping
\cite{bird2009natural}. The language was invented by Guido Van Rossum as a
multi-use, flexible, interpreted scripting language, extensible via compiled C
modules\cite{An93pythonfor}. Developing Partridge in python
would provide a flexible prototyping environment and readable syntax allowing
for investigative work to be carried out quickly and efficiently with minimal
overheads for writing boiler-plate code and debugging. The ability to extend
the language using native code would give Partridge an even greater advantage.
Since NLP techniques can be quite processor intensive and Python is an
interpreted language, and in its very nature, slower than a compiled language,
inefficient sections of Python code could be re-written as C extensions and
compiled into the application. For these reasons, Python was selected as the
preferred programming language for Partridge development.

\subsubsection{User Interface Style}

\subsubsection{Web Presentation Frameworks}

\subsubsection{Natural Language Libraries}
\label{sec:libschoice}.
There are several existing libraries to facilitate Natural Language Processing.
Many are written for Java \cite{mallet2002}\cite{cunningham2011text} and are
very complex or not well documented. The Natural Language Toolkit (NLTK) is a
simple and intuitive library written for Python. Bird(2009) states that NLTK
was designed ``to provide an intuitive framework along with substantial
building blocks, giving users a practical knowledge of
NLP\cite{bird2009natural}". The project is
relatively mature in comparison to the above named Java libraries. There is
also a free book that accompanies the project available at
\url{http://nltk.org/book/} which provides a huge amount of information on how
to implement many popular NLP techniques using the library. For this reason
NLTK was chosen as the primary accompanying library for the project.

@@

\subsection{Development Methodology}

\subsubsection{Agile}
\subsubsection{Waterfall Development}

\subsubsection{ Partridge's methodology}

\subsection{ Work Timeline }
